---
title: "Chapter 2 - ISLR Exercises"
date: "October 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Chapter 2, Question 8

Import the data "College.csv", available here:

http://www-bcf.usc.edu/~gareth/ISL/data.html

> (a) Use the read.csv() function to read the data into R. Call the
> loaded data college. Make sure that you have the directory set
> to the correct location for the data.

We use readr::read_csv().

```{r}
college <- read_csv("College.csv")
```

Note the data was missing a column header for college name. read_csv imputed a column name, 'X1'. We can fix this with dplyr::rename().  

```{r}
college <- college %>%
  rename(school = X1) 
```

> Use the summary() function to produce a numerical summary of the variables in the data set.

Base R functions work just fine. summary() is an excellent example.

```{r}
summary(college)
```

Since using readr::read_csv(), we can also just call college to get info on atomic class of each column in the table.

```{r}
college
```

> Use the pairs() function to produce a scatterplot matrix of the first ten columns or variables of the data. 

Scatterplot matrices work best with numeric data. We can select only the numeric columns with dplyr::select_if(). And then select the first ten columns with dplyr::select().

```{r}
college %>%
  select_if(is_numeric) %>% 
  select(1:10) %>%
  pairs()
```

> Use the plot() function to produce side-by-side boxplots of Outstate versus Private.

Inside of ggplot2::ggplot(), we set the x and y-axes with aes(), then add a geom layer to create the boxplots.

```{r}
college %>% 
  ggplot(aes(x = Private, y = Outstate)) + geom_boxplot()
```

> Create a new qualitative variable, called Elite, by binning
> the Top10perc variable. We are going to divide universities
> into two groups based on whether or not the proportion
> of students coming from the top 10 % of their high school
> classes exceeds 50 %.

dplyr::mutate() makes it much easier to create a new 'Elite' variable.

```{r}
college <- college %>% 
  mutate(Elite = if_else(Top10perc > 50, "Yes", "No"))
```

> Use the summary() function to see how many elite universities
there are. 

Or we can use dplyr::count().

```{r}
college %>%
  count(Elite)
```

> Now use the plot() function to produce side-by-side boxplots of Outstate versus Elite.

Same ggplot as above, but this time we first pass the 'Elite' argument into the aes(), to set the aesthetics of our plot. 

```{r}
college %>% 
  ggplot(aes(Elite, Outstate)) + geom_boxplot()
```

Unsurpising discovery that Elite schools appear to have higher tuition, on average. 

> v. Use the hist() function to produce some histograms with
> diﬀering numbers of bins for a few of the quantitative variables.
> You may find the commandpar(mfrow=c(2,2)) useful:
> it will divide the print window into four regions so that four
> plots can be made simultaneously. Modifying the arguments
> to this function will divide the screen in other ways.

Add your variable of interest -- say, Grad.Rate -- to the aes(), and switch the geom layer to geom_histogram().

```{r}
college %>%
  ggplot(aes(Grad.Rate)) + geom_histogram()
```

